### 2024-03-09 docs/manuscript/manuscript.qmd
figuras de dispersión

```{r prep-env}
#| include: false
#| message: false
#| warning: false
library(flextable)
library(data.table)
load("../../data/mindfulness.RData")
```

@@

We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = 0.258, p = 0.027) response during the TSMT. In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020). No other body composition variables were found to influence the cardiac autonomic response to exercise.

![](figures/fig-1.jpeg)
**Figure 1**. Main correlations among psychological, cardiac autonomic and performance variables. Non-significant correlations (i.e., p > 0.05) were crossed with and "X". The more intense the color the stronger the correlation, either positive or negative. GDS, Geriatric Depression Scale; BAI, Beck Anxiety Inventory; SPPB, Short Physical Performance Battery; TMST, Two-Minutes Step Test; BMI, Body Mass Index.

@@

@@

@@

Fig. X. Diagram of main results.

****Insertar Figura resumen resultados****

Footnote: The figure represents the main results of this study. Blue arrows and boxes represent the correlation between variables, while orange arrows and boxes represent the factors mediating the above correlations. "+" means that the mediating factor enhances the main observed effect, contrary to "-", reflecting a diminished main effect when considering the mediating factor.

>>>>

```{r prep-env}
#| include: false
#| message: false
#| warning: false
library(flextable)
library(data.table)
library(ggplot2)
load("../../data/mindfulness.RData")
dat <- mindfulness[id_event_name == "Basal", -2L]
```

@@

We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = -0.258, p = 0.027) response during the TSMT. In addition, body fat percentage was positively correlated with time domain HRV during exercise, more specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020). No other body composition variables were found to influence the cardiac autonomic response to exercise.

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 4
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "cc_grasa_total_porcentaje"), 
  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
)

dat_long[, variable := fcase(
  grepl("sns", variable), "SNS index",
  grepl("pns", variable), "PNS index",
  grepl("stress", variable), "Stress index",
  grepl("sdnn", variable), "SDNN",
  grepl("rmssd", variable), "RMSSD"
)]

dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]

ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    cc_grasa_total_porcentaje = c(61, 61, 61, 61, 61),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.029", "p = 0.020", "p = 0.262", "p = 0.030", "p = 0.026")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Body fat (%)", x = "Values") +
  tidybayes::theme_ggdist()
```

**Figure 1**. Bivariate dispersion plots between body fat and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.

@@

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 4
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "cc_grasa_total_porcentaje", "cc_masa_muscular_total"), 
  measure.vars = c("tm_pasos", "sppb_total")
)

dat_long[, variable := fcase(
  grepl("tm_pasos", variable), "TMST steps",
  grepl("sppb_total", variable), "SPPB total score"
)]

dat_long[, variable := factor(variable, levels = c("TMST steps", "SPPB total score"))]

fig2a <- ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("TMST steps", "SPPB total score")),
    cc_grasa_total_porcentaje = c(60, 60),
    value = c(42, 4),
    text = c("p = 0.013", "p = 0.015")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Body fat (%)", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

fig2b <- ggplot(dat_long, aes(value, cc_masa_muscular_total)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("TMST steps", "SPPB total score")),
    cc_masa_muscular_total = c(65.5, 65.5),
    value = c(42, 4),
    text = c("p = 0.541", "p = 0.584")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Muscle mass (kg)", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

ggpubr::ggarrange(
  fig2a, fig2b, nrow = 1
)
```
**Figure 2**. Bivariate dispersion plots between physical fitness and body composition related measures. Significance values for the Pearson's product-moment correlation test is shown.

@@

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 8
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "tm_pasos", "sppb_total"), 
  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
)

dat_long[, variable := fcase(
  grepl("sns", variable), "SNS index",
  grepl("pns", variable), "PNS index",
  grepl("stress", variable), "Stress index",
  grepl("sdnn", variable), "SDNN",
  grepl("rmssd", variable), "RMSSD"
)]

dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]

fig3a <- ggplot(dat_long, aes(value, tm_pasos)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    tm_pasos = c(125, 125, 125, 125, 125),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.014", "p = 0.141", "p = 0.002", "p = 0.022", "p = 0.261")
  ), size = 3.5, fontface = "italic") +
  labs(y = "TMST steps", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist() +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.x = element_blank())

fig3b <- ggplot(dat_long, aes(value, sppb_total)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    sppb_total = c(14, 14, 14, 14, 14),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.010", "p = 0.419", "p < 0.001", "p = 0.006", "p = 0.095")
  ), size = 3.5, fontface = "italic") +
  labs(y = "SPPB score", x = "Values") +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

ggpubr::ggarrange(
  fig3a, fig3b, ncol = 1
)
```
**Figure 3**. Bivariate dispersion plots between physical fitness and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.

@@

![](figures/fig_esquema.jpg)
**Figure 4**. Diagram of main results observed form the data. Blue arrows and boxes represent the correlation between variables, while orange arrows and boxes represent the factors mediating the above correlations. "+" means that the mediating factor enhances the main observed effect, contrary to "-", reflecting a diminished main effect when considering the mediating factor.
